---
title: "p8105_hw1_jc6422"
author: "Jianing Chen"
date: "2024-09-18"
output: github_document
---
## Problem 1

load the libraries

```{r}
library(palmerpenguins)
library(ggplot2)
library(tidyverse)
```

load the dataset
```{r}
data("penguins", package = "palmerpenguins")
```

Description of the dataset

```{r}
n_rows = nrow(penguins)
n_cols = ncol(penguins)
mean_flipper_length = mean(penguins$flipper_length_mm, na.rm = TRUE)
cat("The penguins dataset has", n_rows, "rows and", n_cols, "columns.\n")
cat("The mean flipper length (in mm) is", round(mean_flipper_length, 2), ".\n")

```
View the dataset.
```{r}
View(penguins)
head(penguins)
```

There are three species of the penguin: Adelie, Chinstrap, Gentoo.
There are three islands where the penguin was observed: Biscoe, Dream, Torgersen.
The dataset includes several variables such as bill length, bill depth, flipper length, body mass, and sex.



Create scatterplot 
```{r}
scatter_plot = ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point(na.rm = TRUE) +
  labs(title = "Scatterplot of Flipper Length vs Bill Length",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)",
       color = "Species")
print(scatter_plot)
```

Save the scatterplot to a file.
```{r}
ggsave("scatterplot_penguins.png", plot = scatter_plot)
```

## Problem 2

Use tibble function and create data frame with different variables.
```{r}
df2 = tibble(
numeric_var = rnorm(10),

logical_var = numeric_var > 0,

character_var = letters[1:10],

factor_var = factor(rep(c("Level1", "Level2", "Level3"), length.out = 10))
)


```


Take the mean of each variable in my dataframe by pull function.

```{r}
mean_numeric = mean(pull(df2, numeric_var)) 
mean_logical = mean(pull(df2, logical_var))
mean_character = mean(pull(df2, character_var))
mean_factor = mean(pull(df2, factor_var)) 

cat("Mean of numeric variable:", mean_numeric, "\n")
cat("Mean of logical variable:", mean_logical, "\n")
cat("Mean of character variable:", mean_character, "\n")
cat("Mean of factor variable:", mean_factor, "\n")
```

Explanation: The mean of numeric variable and logical variable works, but the mean of character variable and factor variable does not.

Convert variables from one type to another.
```{r}
numeric_logical = as.numeric(df2$logical_var)
numeric_character = as.numeric(df2$character_var)
numeric_factor = as.numeric(df2$factor_var)

mean_numeric_logical = mean(numeric_logical)
mean_numeric_character = mean(numeric_character)
mean_numeric_factor = mean(numeric_factor)

cat("Mean of converted logical variable:", mean_numeric_logical, "\n")
cat("Mean of converted character variable:", mean_numeric_character, "\n")
cat("Mean of converted factor variable:", mean_numeric_factor, "\n")
```
Explanation: The mean of converted logical variable and converted factor variable works, but converted character variable does not work.

Numeric variables are straightforward and can find the mean by arithmetic operations.
Logical variables are coerced to 1 and 0 respectively in arithmetic operations.
Character variables does not represent numbers and thus are not able to have arithmetic operations.
Factor variables are ordered or unordered categorical variable, it cannot directly have arithmetic operations but can have arithmetic operations by coercing a factor to a numeric type using as.numeric() because R treats factors as integers based on the levels, but this is typically not meaningful unless explicitly intended.



