---
title: "p8105_hw1_jc6422"
author: "Jianing Chen"
date: "2024-09-18"
output: github_document
---
## Problem 1

load the libraries

```{r}
library(palmerpenguins)
library(ggplot2)
library(tidyverse)
```

load the dataset
```{r}
data("penguins", package = "palmerpenguins")
```

Description of the dataset

```{r}
n_rows = nrow(penguins)
n_cols = ncol(penguins)
mean_flipper_length = mean(penguins$flipper_length_mm, na.rm = TRUE)
cat("The penguins dataset has", n_rows, "rows and", n_cols, "columns.\n")
cat("The mean flipper length (in mm) is", round(mean_flipper_length, 2), ".\n")

```
View the first few rows of the dataset.
```{r}
head(penguins)
```

Create scatterplot 
```{r}
scatter_plot <- ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() +
  labs(title = "Scatterplot of Flipper Length vs Bill Length",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)",
       color = "Species")
print(scatter_plot)
```

Save the scatterplot to a file.
```{r}
ggsave("scatterplot_penguins.png", plot = scatter_plot)
```

## Problem 2

Create data frame with different variables.
```{r}
# Create a random sample of size 10 from a standard normal distribution
numeric_var = rnorm(10)

# Logical vector indicating whether elements of the sample are greater than 0
logical_var = numeric_var > 0

# Character vector of length 10
character_var = letters[1:10]

# Factor vector of length 10, with 3 different factor levels
factor_var = factor(rep(c("Level1", "Level2", "Level3"), length.out = 10))

df2 = data.frame(numeric_var, logical_var, character_var, factor_var)
print(df2)

```
Take the mean of each variable in my dataframe.

```{r}
mean_numeric <- mean(pull(df, numeric_var)) 
mean_logical <- mean(pull(df, logical_var))
mean_character <- tryCatch(mean(pull(df, character_var)), error = function(e) e)
mean_factor <- tryCatch(mean(pull(df, factor_var)), error = function(e) e) 

cat("Mean of numeric variable:", mean_numeric, "\n")
cat("Mean of logical variable:", mean_logical, "\n")
cat("Mean of character variable:", mean_character$message, "\n")
cat("Mean of factor variable:", mean_factor$message, "\n")

```

The mean of numeric variable and logical variable works, but the mean of character variable and factor variable does not.


